---
title: "Multiple Linear Regression Stat Project 2"
output: html_document
date: "2023-12-06"
---
# TODO WYATT NOTES:

#1. Check for outliers and high impact variables (one house has 33 bedrooms)
#2. CHECK FOR MULTICOLLINEARITY

```{r}
library(tidyverse)

setwd('C:/Users/wyatt/OneDrive/Documents/repos/stat6021-project2')

# Read in Data
data <- read.csv('kc_house_data.csv')

# View headers
head(data, 3)
```

```{r}
colnames(data)
```
# Data Definitions
<!-- id - Unique ID for each home sold --> DON'T WANT FOR ANALYSIS

<!-- date - Date of the home sale --> Only have 2 years - should not have a big impact on price of houses within the market with this small of a time frame
```{r}
years_sold <- data$date %>%
  substr(1, 4)

table(years_sold)
```

<!-- price - Price of each home sold -->
# RESPONSE VARIABLE ^ NEED :)
# Potentiall take price in '000s
```{r}
summary(data$price/1000)
```
<!-- bedrooms - Number of bedrooms -->
# KEEP IN REGRESSION
```{r}
table(data$bedrooms)
```
<!-- bathrooms - Number of bathrooms, where .5 accounts for a room with a toilet but no shower -->
# KEEP IN REGRESSION
```{r}
summary(data$bathrooms)
```
<!-- sqft_living - Square footage of the apartments interior living space -->
# POTENTIALLY KEEP IN REGRESSION
```{r}
summary(data$sqft_living)
```
<!-- sqft_lot  - Square footage of the land space -->
# POTENTIALLY KEEP IN REGRESSION
```{r}
summary(data$sqft_lot)
```
<!-- sqft_above - The square footage of the interior housing space that is above ground level -->
```{r}
summary(data$sqft_above)
```
<!-- sqft_basement - The square footage of the interior housing space that is below ground level -->
```{r}
summary(data$sqft_basement)
```

<!-- floors - Number of floors -->
#Potentially keep in regression - probably not though
```{r}
table(data$floors)
```
<!-- waterfront - A dummy variable for whether the apartment was overlooking the waterfront or not -->
# proabbly drop,not big sample size difference - just keep view
```{r}
table(data$waterfront)
```
<!-- view - An index from 0 to 4 of how good the view of the property was -->
# Maybe just keep view though
```{r}
table(data$view)
```
<!-- condition - An index from 1 to 5 on the condition of the apartment, -->
# definately keep condition
```{r}
table(data$condition)
```
<!-- grade - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design. -->
# potentially keep grade
```{r}
table(data$grade)
```
<!-- yr_built - The year the house was initially built -->
# Potentially Keep
```{r}
table(data$yr_built)
```
<!-- yr_renovated - The year of the house’s last renovation -->
# Probably don't keep year renovated OR convert to dummy variable for Renovated or Not
```{r}
table(data$yr_renovated)
```
# NEIGHBORHOOD VARIABLES
<!-- zipcode - What zipcode area the house is in -->
```{r}
table(data$zipcode)
```
<!-- lat - Latitude -->
# DONT NEED LAT OR LONG
<!-- long - Longitude -->
# DONT NEED LAT OR LONG

<!-- sqft_living15 - The square footage of interior housing living space for the nearest 15 neighbors -->
```{r}
summary(data$sqft_living15)
```
<!-- sqft_lot15 - The square footage of the land lots of the nearest 15 neighbors -->

```{r}
summary(data$sqft_lot15)
```
