---
title: "Multiple Linear Regression Stat Project 2"
output: html_document
date: "2023-12-06"
---
# TODO WYATT NOTES:

#1. Check for outliers and high impact variables (one house has 33 bedrooms)
#2. Review why logging variables
#3. use automated search procedures for model from existing vars

```{r}
library(tidyverse)
library(MASS)
library(leaps)
library(faraway)

#Disable scientific notation
options(scipen = 100)


setwd('C:/Users/wyatt/OneDrive/Documents/repos/stat6021-project2')

# Read in Data
data <- read.csv('kc_house_data.csv')


# Remove Data with no bedrooms or bathrooms
data <- subset(data, data$bathroom != 0 | data$bedrooms != 0)

# View headers
head(data, 3)
```

```{r}
colnames(data)
```

# Data Definitions AND Summary Stats
<!-- id - Unique ID for each home sold --> DON'T WANT FOR ANALYSIS

<!-- date - Date of the home sale --> Only have 2 years - should not have a big impact on price of houses within the market with this small of a time frame
```{r}
years_sold <- data$date %>%
  substr(1, 4)

table(years_sold)
```

<!-- price - Price of each home sold -->
# RESPONSE VARIABLE ^ NEED :)
# Potentiall take price in '000s
```{r}
summary(data$price/1000)
```
<!-- bedrooms - Number of bedrooms -->
# KEEP IN REGRESSION
```{r}
table(data$bedrooms)
```
<!-- bathrooms - Number of bathrooms, where .5 accounts for a room with a toilet but no shower -->
# KEEP IN REGRESSION
```{r}
table(data$bathrooms)
```
<!-- sqft_living - Square footage of the apartments interior living space -->
# POTENTIALLY KEEP IN REGRESSION
```{r}
summary(data$sqft_living)
```
<!-- sqft_lot  - Square footage of the land space -->
# POTENTIALLY KEEP IN REGRESSION
```{r}
summary(data$sqft_lot)
```
<!-- sqft_above - The square footage of the interior housing space that is above ground level -->
```{r}
summary(data$sqft_above)
```
<!-- sqft_basement - The square footage of the interior housing space that is below ground level -->
```{r}
summary(data$sqft_basement)
```

<!-- floors - Number of floors -->
#Potentially keep in regression - probably not though
```{r}
table(data$floors)
```
<!-- waterfront - A dummy variable for whether the apartment was overlooking the waterfront or not -->
# proabbly drop,not big sample size difference - just keep view
```{r}
table(data$waterfront)
```
<!-- view - An index from 0 to 4 of how good the view of the property was -->
# Maybe just keep view though
```{r}
table(data$view)
```
<!-- condition - An index from 1 to 5 on the condition of the apartment, -->
# definately keep condition
```{r}
table(data$condition)
```
<!-- grade - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design. -->
# potentially keep grade
## CREATEINTO BUCKETS
```{r}
table(data$grade)
```
<!-- yr_built - The year the house was initially built -->
# Potentially Keep
```{r}
table(data$yr_built)
```
<!-- yr_renovated - The year of the house’s last renovation -->
# Probably don't keep year renovated OR convert to dummy variable for Renovated or Not
```{r}
table(data$yr_renovated)
```
# NEIGHBORHOOD VARIABLES
<!-- zipcode - What zipcode area the house is in -->
# Geo spatial regression too complicated for this analysis
```{r}
table(data$zipcode)
```
<!-- lat - Latitude -->
# DONT NEED LAT OR LONG
<!-- long - Longitude -->
# DONT NEED LAT OR LONG

<!-- sqft_living15 - The square footage of interior housing living space for the nearest 15 neighbors -->
```{r}
summary(data$sqft_living15)
```
<!-- sqft_lot15 - The square footage of the land lots of the nearest 15 neighbors -->

```{r}
summary(data$sqft_lot15)
```
## SUBSECT INITIAL DATA REVIEW

# Create new column for renovated boolean and grade category
```{r}
# Add renovated boolean
data$renovated <- as.factor(ifelse(data$yr_renovated == 0, 0,1))

# Create Quality buckets FROM GRADE ##NEW VARIABLE
data$grade <- cut(as.numeric(data$grade), breaks = c(0, 3, 10, 14), right=FALSE, labels = c('subpar', 'average', 'high quality'))
data$grade <- relevel(data$grade, "average")


#Drop ID - ID only differentiates rows so not needed
# Date - there are only two years of data here so not much variation for time series
# Zipcode - omit regional data to simplify regression
# lat, long -omit specific location to simplify regression
# sqft_basement - omit since NA - linearly related to other variables
#TRANSFORM GRADE, yr_built, #TRANSFORM yr Renovated, 
data <- data %>%
             dplyr::select(price, 
                           bedrooms, 
                           bathrooms, 
                           sqft_living,
                           sqft_lot,
                           floors,
                           waterfront,
                           view, 
                           condition,
                           grade, 
                           yr_built,
                           sqft_above,
                           renovated,
                           sqft_living15,
                           sqft_lot15
                           )

# Create TEST/TRAIN Split
set.seed(6021)
sample.data<-sample.int(nrow(data), floor(.50*nrow(data)), replace = F)


train<-data[sample.data, ]
rownames(train) <- NULL
test<-data[-sample.data, ]
rownames(test) <- NULL

# Get correlation coefficients
round(cor(train[,unlist(lapply(train, is.numeric), use.names = FALSE)]), 3)


# start visualization 
pairs(train[,unlist(lapply(train, is.numeric), use.names = FALSE)])
# Floors, waterfront, views, and conditions are categorical variables
```


# Convert Categorical Variables to factors
```{r}
train$floors <- as.factor(train$floors)
train$view <- as.factor(train$view)
train$condition <- as.factor(train$condition)
train$waterfront <- as.factor(train$waterfront)
```


```{r}
# Run initial regression
# Looking at price in '000s
results <- lm(price/1000~., data=train)
summary(results)
```

# MULTICOLLinearity
# sqft living and sqft above both indicate multicollinearity. drop sqft_above as sqft_living is more relevant to home price. condition has high multi-collinearity, drop from model
```{r}
faraway::vif(results)
```
# Re-run regression w/o sqft_above and condition
```{r}
train <- train %>%
                     dplyr::select(price, 
                                   bedrooms, 
                                   bathrooms, 
                                   sqft_living,
                                   sqft_lot,
                                   floors,
                                   waterfront,
                                   view, 
                                   grade, 
                                   yr_built,
                                   renovated,
                                   sqft_living15,
                                   sqft_lot15
                                   )

# Run initial regression
results <- lm(price/1000~., data=train)
summary(results)
```
#no longer evidence of multicollinearity

```{r}
faraway::vif(results)
```

# Automated search procedures

# utilizing stepwise prediction 
```{r}

reg_null <- lm(price/1000~1, data=train)

step(reg_null, scope=list(lower=reg_null, upper=results), direction='both')
```

```{r}
# Build model based on results of step-wise automated search procedures
automated_results <- lm(formula = price/1000 ~ sqft_living + view + grade + yr_built + 
    floors + waterfront + sqft_living15 + bathrooms + bedrooms + 
    sqft_lot15, data = train)

summary(automated_results)
```

# Assess assumptions

```{r}
#assess regression assumptions
par(mfrow=c(2,2))
plot(automated_results)

# ASSUMPTION 2 are not met so we need to transform the response variable (price)
```


```{r}
# 1 does not lie in the CI AND there is increasing variance from left to right therefore a transformation of Y may be needed - we will use log so we can still interpret the estimated coefficients
boxcox(automated_results)
```

# Re-run regression with transformed Y
```{r}
transformed_results <- lm(formula = log(price) ~ sqft_living + view + grade + yr_built + 
    floors + waterfront + sqft_living15 + bathrooms + bedrooms + 
    sqft_lot15, data = train)



# Example interpretation:
#1.For a 1 unit increase in sqft living, there is a 0.022% increase in price accounting for all other variables
#2. For a 1 unit increase in year built, there is a 0.44% decrease in price accounting for all other variables
summary(transformed_results)
```

# Assumptions look much better - go with this model
```{r}
plot(transformed_results)
```




```
```{r}
#confidence intervals for coefficients
confint(transformed_results, level=0.95)
```


