---
title: "Multiple Linear Regression Stat Project 2"
output: html_document
date: "2023-12-06"
---
# TODO WYATT NOTES:

#1. Check for outliers and high impact variables (one house has 33 bedrooms)
#2. Review why logging variables
#3. use automated search procedures for model from existing vars

```{r}
library(tidyverse)
library(MASS)
library(GGally)
library(leaps)
library(faraway)

#Disable scientific notation
options(scipen = 100)


# Read in Data
data <- read.csv('C:/Users/wyatt/OneDrive/Documents/repos/stat6021-project2/kc_house_data.csv')


# Remove Data with no bedrooms or bathrooms
data <- subset(data, data$bathroom != 0 | data$bedrooms != 0)

# Fix data entry error
data[which(data$bedrooms == 33),'bedrooms'] <- 3

# View headers
head(data, 3)
```

```{r}
colnames(data)
```

# Data Definitions AND Summary Stats
# Create new column for renovated boolean and grade category
```{r}
# Add renovated boolean
data$renovated <- as.factor(ifelse(data$yr_renovated == 0, 0,1))

# Create Quality buckets FROM GRADE ##NEW VARIABLE
data$grade <- cut(as.numeric(data$grade), breaks = c(0, 3, 10, 14), right=FALSE, labels = c('subpar', 'average', 'high quality'))
data$grade <- relevel(data$grade, "average")


#Drop ID - ID only differentiates rows so not needed
# Date - there are only two years of data here so not much variation for time series
# Zipcode - omit regional data to simplify regression
# lat, long -omit specific location to simplify regression
# sqft_basement - omit since NA - linearly related to other variables
#TRANSFORM GRADE, yr_built, #TRANSFORM yr Renovated, 
data <- data %>%
             dplyr::select(price, 
                           bedrooms, 
                           bathrooms, 
                           sqft_living,
                           sqft_lot,
                           floors,
                           waterfront,
                           view, 
                           condition,
                           grade, 
                           yr_built,
                           sqft_above,
                           renovated,
                           sqft_living15,
                           sqft_lot15
                           )

# Create TEST/TRAIN Split
set.seed(6021)
sample.data<-sample.int(nrow(data), floor(.50*nrow(data)), replace = F)


# Floors, waterfront, views, and conditions are categorical variables
# Convert Categorical Variables to factors

data$floors <- as.factor(data$floors)
data$view <- as.factor(data$view)
data$condition <- as.factor(data$condition)
data$waterfront <- as.factor(data$waterfront)

train<-data[sample.data, ]
rownames(train) <- NULL
test<-data[-sample.data, ]
rownames(test) <- NULL


# Get correlation coefficients and Matrix

GGally::ggpairs(train[,unlist(lapply(train, is.numeric), use.names = FALSE)], progress = FALSE,upper=list(continuous = wrap("cor", size=2))) +
  theme(axis.text = element_text(size = 5),
        strip.text.y = element_text(size = 8,angle = 0),
        strip.text.x = element_text(size = 8,angle = 90),
        legend.position = "none",
        panel.grid.major = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_text(angle = 90, vjust = 1, color = "black"),
        panel.border = element_rect(fill = NA))
```



```


```{r}
# Run initial regression
# Looking at price in '000s
results <- lm(price/1000~., data=train)
summary(results)
```

# MULTICOLLinearity
# sqft living and sqft above both indicate multicollinearity. drop sqft_above as sqft_living is more relevant to home price. condition has high multi-collinearity, drop from model
```{r}
faraway::vif(results)
```
# Re-run regression w/o sqft_above and condition
```{r}
train <- train %>%
                     dplyr::select(price, 
                                   bedrooms, 
                                   bathrooms, 
                                   sqft_living,
                                   sqft_lot,
                                   floors,
                                   waterfront,
                                   view, 
                                   grade, 
                                   yr_built,
                                   renovated,
                                   sqft_living15,
                                   sqft_lot15
                                   )
                                   
# Do same for test results
test <- test %>%
                     dplyr::select(price, 
                                   bedrooms, 
                                   bathrooms, 
                                   sqft_living,
                                   sqft_lot,
                                   floors,
                                   waterfront,
                                   view, 
                                   grade, 
                                   yr_built,
                                   renovated,
                                   sqft_living15,
                                   sqft_lot15
                                   )

# Run initial regression
results <- lm(price/1000~., data=train)
summary(results)
```
#no longer evidence of multicollinearity

```{r}
faraway::vif(results)
```

# Automated search procedures

# utilizing stepwise prediction 
```{r}

reg_null <- lm(price/1000~1, data=train)

step(reg_null, scope=list(lower=reg_null, upper=results), direction='both')
```

```{r}
# Build model based on results of step-wise automated search procedures
automated_results <- lm(formula = price/1000 ~ sqft_living + view + grade + yr_built + 
    floors + waterfront + sqft_living15 + bathrooms + bedrooms + 
    sqft_lot15, data = train)

summary(automated_results)
```

# Assess assumptions

```{r}
#assess regression assumptions
par(mfrow=c(2,2))
plot(automated_results)

# ASSUMPTION 2 are not met so we need to transform the response variable (price)
```


```{r}
# 1 does not lie in the CI AND there is increasing variance from left to right therefore a transformation of Y may be needed - we will use log so we can still interpret the estimated coefficients
boxcox(automated_results)
```

# Re-run regression with transformed Y
```{r}
transformed_results <- lm(formula = log(price) ~ sqft_living + view + grade + yr_built + 
    floors + waterfront + sqft_living15 + bathrooms + bedrooms + 
    sqft_lot15, data = train)



# Example interpretation:
#1.For a 1 unit increase in sqft living, there is a 0.022% increase in price accounting for all other variables
#2. For a 1 unit increase in year built, there is a 0.44% decrease in price accounting for all other variables
summary(transformed_results)
```

# Assumptions look much better - go with this model
```{r}
plot(transformed_results)
```




```
```{r}
#confidence intervals for coefficients
confint(transformed_results, level=0.95)
```

```{r}
test_predictions <- predict(transformed_results, newdata = test)
test_rmse <- sqrt(mean((log(test$price) - test_predictions)^2))

# Relevant Interpretations
interpretation <- summary(transformed_results)

```


